\LARGE \bf Mean Estimation from Adaptive Single-bit Measurements

### 2017-09-26 Writeups/sketch_distributed_scaling.tex
Add files via upload

%\documentclass[letterpaper]{IEEEtran}
\documentclass[letterpaper, conference]{IEEEtran}      % Use this line for a4 paper

%LatexiDiff ignore tikz:
% $: latexdiff -c "PICTUREENV=(?:picture|tikzpicture|DIFnomarkup)[\w\d*@]*" old.tex new.tex > diff.tex

%\IEEEoverridecommandlockouts                              %

@@

\title{\LARGE \bf Mean Estimation from Single-bit Measurements}
%
%\author{
%\IEEEauthorblockN{Alon Kipnis}
%\IEEEauthorblockA{Department of Electrical Engineering \\
%Stanford University\\
%Stanford, CA\\}
%\and
%\IEEEauthorblockN{John C. Duchi}
%\IEEEauthorblockA{Department of Statistics \\
%and Department of Electrical Engineering \\
%Stanford University\\
%Stanford, CA\\}
%\and
%\IEEEauthorblockN{Andrea J. Goldsmith}
%\IEEEauthorblockA{Department of Electrical Engineering \\
%Stanford University\\
%Stanford, CA\\}
%}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
We consider the problem of estimating the mean of a normal distribution under the constraint that only a single bit per sample from this distribution is available to the estimator. We study the mean squared error risk in this estimation as a function of the number of samples, and hence number of bits, from this distribution. We consider an adaptive scheme in which each bit is a function of the current sample and the previously observed bits, as well as a fully distributed setting in which each bit is only a function of the current sample. For the adaptive scheme, we show that the optimal relative efficiency, compared to the standard mean estimator without bit limitation, is $\pi/2$. Namely, the number of samples required to attain the mean to a prescribed accuracy is only $\pi/2$ times larger due to the bit constraint. For the distributed scheme we consider the setting where each bit is obtained by comparing the current sample to a prescribed value. For this setting, we show that the expected asymptotic efficiency converges to $\pi/2$ provided that the variance of the distribution is high compared to the apriori uncertainty about the mean. Our results indicate that, rather surprisingly, parametric estimation from the coarsest form of quantization without significant penalty in the number of measurements is possible. 
\end{abstract}

%{\color{red}  The $\mu$-sum problem (Kim-ElGamal Ch. 21) gives a lower bound. Can be reduced to the CEO.}

%{\color{red}  See LASSO results in http://arxiv.org/abs/1506.02181v1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Background}
\label{sec:Intro}
Processing information on multiple physical locations may impose communication constraints on the amount of information that can be shared among different units of the system. 
%For example, emerging sensor network technologies in medicine and ``smart cities" use many low-cost sensors to collect large amounts of data and transmit it to remote locations. These sensors must operate under severe power restrictions, hence they are limited by their communication bandwidth. 
%For example, this constraint occurs in large-scale sensor arrays where information is collected at multiple physical locations and transmitted to a central estimation unit. 
In this situation, the ability to estimate a particular parameter is dictated not only by quality of observations but also by the constraint on the communication rate of the sensor, which may be severely restricted due to power, bandwidth, or size of data. The question that we ask is to what extent a parametric estimation task is affected by such communication constraints, and what are the fundamental performance of estimation subject to these restrictions. \\

In this paper we provide a precise answer to this question in a limited setting: the estimation of the mean $\theta$ of a normal distribution with a known variance $\sigma^2$, under the constraint that only a single bit can be communicated to the estimator on each sample from this distribution. As it turns out, the ability to share information before communicating it dramatically affects the performance in recovering $\theta$. We distinguish in particular among three different settings:
 \begin{itemize}
 \item[(i)] \emph{Centralized} encoding: all the $n$ encoders confer and produce a single $n$ bit message. 
 \item[(ii)] \emph{Adaptive} or \emph{sequential} encoding: the $i$th encoder observes the $i$th sample and the $i-1$ previously encoded observations.
 \item[(iii)] \emph{Distributed} encoding: the output of the $i$th encoder is only a function of the $i$th observation.
 \end{itemize}
As far as information sharing is concerned, it is evident that settings (iii) is a more restrictive version of (ii) which is more restrictive than (i). We are concerned with the mean squared error (MSE) risk in estimating $\theta$ as the number of observations $n$ goes to infinity. In particular, we are interested in the \emph{relative efficiency} of estimators in settings (i)-(iii), defined as the ratio between their MSE to $\sigma^2/n$. The latter is the MSE attained by the sample mean, which is the minimal risk estimator when no constraints on the communication are imposed. \par
%
It is well known that the relative efficiency in setting (i) is $1$ \cite{720540, zhang2013information, zhang1988estimation}, i.e., asymptotically, there is no loss in performance due to the communication constraint in this setting. Indeed, under setting (i) the sample mean can be described with an error exponentially decaying in $n$, leading to MSE of $\sigma^2/n + o(n)$. In this work we show that a similar result does not hold even in setting (ii): the relative efficiency of any adaptive estimator is at least $\pi/2$. Namely, a single bit per sample communication constraint in the adaptive setting incurs a minimal penalty of at least $\pi/2 \approx 1.57$ compared to an unconstrained estimator, or to the optimal estimator in setting (i). In addition to this negative statement, we provide an estimator which asymptotically attains this minimal relative efficiency, i.e., we show that this lower bound is tight. Moreover, we show that even in the distributed setting (iii), a relative efficiency of $\pi/2$ is possible as long as $\sigma^2$ is large compared to the variance of the prior distribution on $\theta$. These results are quite surprising since they show that, even under the coarsest form of quantization, an estimator based on $ \approx 1.59 n$ measurements achieves accuracy in MSE or confidence interval comparable to an estimator that uses $n$ samples with infinite bit-precision.  \par
%
%One clue for this surprising results is given in a famous problem in information theory denoted as the CEO problem, first presented by Berger et. al. in \cite{berger1996ceo}. In this problem, a firms CEO is required to estimate a signal  from information sent to it by multiple agents, where the $i$th agents is given a noisy observation of this signal and has a limited number of bits, say $R_i$, to describe its observation to the CEO.

\subsection*{Related Works}
%\subsubsection*{Statistical inference under communication constraints}
Statistical inference problems under communication constraints were considered in information theory in 
\cite{han1987hypothesis, zhang1988estimation, 720540}. These works consider a multi-terminal source coding setting in which the $i$th terminal observes $n$ samples from the distribution and is allotted $nR_i$ bits to communicate its estimate. The main focus of these works is the difference between inference with communication constraint and the unconstrained vanilla statistical estimation setting, as the number of samples $n$ goes to infinity subject to a total finite rate constraint. Our setting (i) can be seen as a special case of this setting where with a single terminal and $R_1 = 1$.% whereas our case (iii) is obtained from this setting by taking $n=1$, $R_i=1$, and number of terminals grow with $n$. 
However, as explained in \cite[Sec. III]{720540}, a single terminal in this setting always lead to the unconstrained inference performance. Indeed, when all samples are taken from the same distribution, the \emph{type} of the sample \cite{csiszar1998method} is a sufficient statistics for any inference task, and the latter can be described using a number of codewords polynomial in $n$ regardless of the distribution of the samples. For this reason, attention is given in these works to inference problems involving multiple distributions observed at different locations and hence the results there do not apply to our setting. In fact, with a full access to the sample as in setting (i), the problem of encoding and estimating $\theta$ is reduced to the MSE attained by a scalar quantizer adjusted to the sufficient statistics of the sample. Finally, we note that setting (ii) includes as a special case the sigma-delta modulation (SDM) analog-to-digital conversion scheme with a constant input $\theta$ corrupted by Gaussian noise $Z_i$, as was considered in \cite{53738}. While it was shown there that the output of the modulator converges to the true constant input, the rate of this convergence was not analyzed and cannot be derived directly from the results of \cite{53738}. In fact, a corollary from the results in this paper we conclude that the rate of convergence of a SDM to a constant input signal is at most $\sigma^2\pi/2$ over the number of feedback iterations. Finally, the multiterminal source coding setting denoted as the CEO problem \cite{berger1996ceo} can be seen as a generalization of our setting (iii), where multiple draws of $\theta$ and rounds of $n$ bit communication is permitted. In particular, the minimal CEO distortion with $1$ bit per terminal provides a lower bound on the MSE risk.  \par
%The difference between this setting and ours is that in ours the parameter of interest is not an information source. By assuming a prior distribution on this parameter, the CEO provides a lower bound on the estimation error in the fully distributed setting (iii). This lower bound can be attained if we were to consider the average error in multiple independent realizations of our problem rather than a single realization as we do here. \\
%
%While our focus in this paper is on setting (ii), we note that unlike the above works, we are also interested in the optimal estimation performance for a finite number of observations and not only in their asymptotic behavior. 
%\subsubsection*{Relation to channel coding}
More related to our setting is the work \cite{zhang2013information} that considers the minimax MSE risk in estimating a vector of parameters from $m$ machines, where each machine has access to $n/m$ samples a limited number of bits to describe these samples. The main results of this work are lower bounds on the estimation error as a function of the total number of bits can be sent by each machine. In the terminology of \cite{zhang2013information}, our settings (ii) and (iii) fall under the category of \emph{interactive}, and  \emph{independent} communication protocols, respectively. %In both cases we use $m=n$ (single sample per machine) and communication constraint of a single bit per machine.
However, unlike in \cite{zhang2013information}, we focus on achievable schemes and tight bounds in $n$ for the asymptotic MSE risk, rather than its scaling rate. \\

The rest of this paper is organized as follows: the main problem is defined in Section~\ref{sec:problem}. Our main results are presented and discussed in Sections~ \ref{sec:sequential} and \ref{sec:distributed}. Concluding remarks are given in Section~\ref{sec:conclusions}.

%We also note that the counterpart of setting (iii) in the case of hypothesis testing was considered in \cite{52470}.

%\subsubsection{Estimation from 1-bit measurements}
%Single-bit or binary detectors are the basis for digital technology, and works considers the estimation of parameters from the results of such detectors have been around from the early days of digital computing. The burst of works in compressed sensing in the previous decade also gave raise to multiple 1-bit results \cite{rice_university}. Due to the popularization of MIMO technology and the expected break of mm-wave in wireless communication there have been many works 1-bit detection at the A/D in the receiver \cite{2016arXiv160204193Z}.

\section{Problem Formulation \label{sec:problem}}

\begin{figure}
\begin{center}
\begin{tikzpicture}[node distance=2cm,auto,>=latex]

\node[node distance = 1.5cm] at (0,0) (source2) {$X_1,\ldots,X_n$};
\node[int1, right of = source2, node distance = 2cm] (enc2) {Enc};  
\draw[->,line width = 2pt] (source2) -- (enc2);

\node[above of = source2, node distance = 1cm] (dist) {$X_i \sim {\mathcal N} \left(\theta, \sigma^2 \right)$};

\node[int1, right of = enc2, node distance = 4cm ] (est) {Est};

\draw[->,line width = 2pt] (enc2) -- node[above, xshift = 0cm] (mes2) {$M \in \left\{-1,1\right\}^n$} (est);

\node[right of = est, node distance = 1.5cm] (dest) {$\widehat{\theta}_n$};
%         \node [int1] (dec) [right of=dest, node distance = 1.5cm,  align=center] {\small Dec };
%\node [int1] (enc) [right of = dec, node distance = 3cm]{Enc}; 
%\draw[->,line width=2pt] (dec) -- (dest);
\draw[->, line width=2pt] (est) -- (dest);
\end{tikzpicture}
\end{center}
\caption{\label{fig:centralized} Centralized encoding using one bit per sample on average.}
\end{figure}

@@

\node[below of = source2, node distance = 0.5cm] {$\vdots$};
\node[above of = source3, node distance = 0.6cm] {$\vdots$};
\node[above of = source3, node distance = 1cm] (dist) {$X_i \sim {\mathcal N} \left(\theta, \sigma^2 \right)$};

\node[int1, right of = enc2, node distance = 4cm ] (est) {Est};
\draw[->,line width = 2pt] (enc1) -| node[above, xshift = -2cm] (mes1) {$M_1 \in \left\{-1,1\right\}$} (est);   
\draw[->,line width = 2pt] (enc2) -- node[above, xshift = 0cm] (mes2) {$M_2 \in \left\{-1,1\right\}$} (est);   
\draw[->] (mes1) -- +(0,-0.7) -| (enc2);

\draw[dashed,->] (mes2) -- +(0,-0.7) -| +(-2,-1.2);
\draw[->,line width = 2pt] (enc3) -| (est);

\node[below of = mes2, node distance = 1.5cm] (mes3) {$M_1,\ldots,M_{n-1} $};

\draw[->] (mes3) -- +(0,-0.5) -| (enc3);

\node[right of = est, node distance = 1.5cm] (dest) {$\widehat{\theta}_n$};
%         \node [int1] (dec) [right of=dest, node distance = 1.5cm,  align=center] {\small Dec };
%\node [int1] (enc) [right of = dec, node distance = 3cm]{Enc}; 
%\draw[->,line width=2pt] (dec) -- (dest);
\draw[->, line width=2pt] (est) -- (dest);
\end{tikzpicture}
\end{center}
\caption{\label{fig:sequential} Adaptive single-bit encoding: the $i$th encoder delivers a single bit message which is a function of its private sample $X_i$ and the previous messages $M_1,\ldots,M_{i-1}$.}
\end{figure}

\node[above of = source3, node distance = 1cm] (dist) {$X_i \sim {\mathcal N} \left(\theta, \sigma^2 \right)$};
\node[below of = source2, node distance = 0.5cm] {$\vdots$};
\node[above of = source3, node distance = 0.6cm] {$\vdots$};

\node[int1, right of = enc2, node distance = 4cm ] (est) {Est};
\draw[->,line width = 2pt] (enc1) -| node[above, xshift = -2cm] (mes1) {$M_1 \in \left\{-1,1\right\}$} (est);

\draw[->,line width = 2pt] (enc2) -- node[above, xshift = 0cm] (mes2) {$M_2 \in \left\{-1,1\right\}$} (est);

@@

\node[right of = est, node distance = 1.5cm] (dest) {$\widehat{\theta}_n$};
%         \node [int1] (dec) [right of=dest, node distance = 1.5cm,  align=center] {\small Dec };
%\node [int1] (enc) [right of = dec, node distance = 3cm]{Enc}; 
%\draw[->,line width=2pt] (dec) -- (dest);
\draw[->, line width=2pt] (est) -- (dest);
\end{tikzpicture}
\end{center}
\caption{\label{fig:distributed} Distributed single-bit encoding: the single-bit message produced by each encoder is only a function of its private sample $X_i$.}
\end{figure}

Let $X_i$, $i=1,\ldots,n$, be $n$ independent samples from the normal distribution $P(X) = \mathcal N(\theta,\sigma^2)$ with mean $\theta$ and variance $\sigma^2$. We also assume that the mean $\theta$ is drawn once from a prior distribution $\pi(\theta)$ on the parameter space $\Theta$, which is a closed interval of the real line. We moreover assume that $\pi(\theta)$ is absolutely continuous with respect to the Lebesgue measure with a bounded second moment. We denote the variance of $\theta$ with respect to $\pi(\theta)$ by $\sigma_\theta^2$. The problem we consider is the estimation of the parameter $\theta$ under the following constraints on the communication between the samples $X_1,\ldots,X_n$ and a centralized estimator: 
\begin{itemize}
\item[(i)] The estimator at time $n$ is only a function of the $n$ messages $M^n = \left(M_1,\ldots,M_n \right)$ (Fig.~\ref{fig:centralized}).
\item[(ii)] For each $i=1,\ldots,n$, the $i$th message $M_i$ is a function of the sample $X_i$ and the $i-1$ previous messages $M^{i-1}$ (Fig.~\ref{fig:sequential}).
\item[(iii)] The $i$th message $M_i$ takes only two possible values, say $1$ and $-1$ (Fig.~\ref{fig:distributed}).
\end{itemize}
In other words, the only information on the sample $X^n \left(X_1,\ldots,X_n\right)$ available to the estimator is the messages $M^n$. In addition, each message $M_i$ is a function from the real line into the set $\{-1,1\}$ which is measurable with respect to the sigma algebra generated by $M^{i-1}$ and $X_i$. The estimator, upon observing $M^n$, produces an estimate $\widehat{\theta}_n(M^n)$ of $\theta$. A system describing the above scheme is illustrated in Fig.~\ref{fig:sequential}. \\

The performance of this estimation is measured by the mean squared error (MSE) risk function:
\begin{equation}
\label{eq:error_def}
R_n \triangleq \mathbb E\left(\widehat{\theta}_n - \theta \right)^2,
\end{equation}
where the expectation is taken with respect to the distribution of $X^n$ and the prior distribution $\pi(\theta)$.  \\
%It is well known that minmax estimation error 
%\begin{equation}
%R_n = \min \sup_{\theta \in \Theta}  \mathbb E\left[ \left(\widehat{\theta}_n - \theta \right)^2 | \theta \right],
%\end{equation}
%can be obtained from $R_n$ by assuming a particular least favorable prior distribution on the parameter space $\Theta$ \cite{casella1981}.

The main problem we consider is the minimal value that can be attain in \eqref{eq:error_def} as a function of $n$. Note that this minimization is the combination of the following two processes: (1) selecting the $i$th message $M_i$ based on past messages and current observation $X_i$, and (2) estimating $\theta$ given messages $M^n$. We are interested in particular in the asymptotic relative efficiency of estimators $\widehat{\theta}_n$ for $\theta$ compared to the sample mean $\bar{X}_n = \frac{1}{n} \sum_{i=1}^n X_i$. Since the MSE attained by the latter is $\sigma^2/n$, this relative efficiency is defined as
\begin{equation}
 \lim_{n \rightarrow \infty} \frac{n}{\sigma^2}  R_n. 
\label{eq:relative_efficiency}
\end{equation}
%{\color{red} Since the conditional expectation of $\theta$ with respect to $\bar{X}_n$ (and not $\bar{X}_n$) minimizes \eqref{eq:error_def}, we need to justify the fact that we use the sample mean as the reference estimator.}

In addition to the notations defined above, we denote by $\phi(x)$ the standard normal density and by $\Phi(x)$ the standard normal cumulative distribution function. Prime denote derivative with respect to $\theta$.

\section{Distributed Estimation \label{sec:distributed}}
We now consider the case where a single bit message is available on each sample from the distribution, and this message is independent of the other messages. Moreover, we assume that each message is characterized by a threshold value, and that ``$1$'' is sent by the $i$th encoder whenever $x_i$ is above this value.

\subsection{Threshold Detection}
We assume that each message is of the form
\[
M_i = \sgn(t_i - X_i) = \begin{cases} 1 & X_i< t_i, \\
-1 & X_i > t_i,
\end{cases}  
\]
where $t_i\in\mathbb R$ is the \emph{threshold} of the $i$th detector applied by the encoder on it sample $X_i$. In order to characterize the asymptotic behavior of estimation from these sequence of messages, we consider the \emph{density} of $n$ threshold values defined as:
\[
\lambda_n([a,b]) = \frac{1}{n} \left| \mathcal T \cap [a,b] \right|,
\]
where $\mathcal T  = \left\{t_n, n \in \mathbb N \right\}$ is the collection of all threshold values. We further assume that $\lambda_n$ converges (weakly) to a probability measure $\lambda(t)$ on $\mathbb R$.  \par
In order to estimate $\theta$ from the messages $M^n$, we consider the maximum likelihood (ML) estimator. The log-likelihood function of $\theta$ is given by
\[
l(M^n  |\theta ) = \sigma  \sum_{i=1}^n \log \left( \Phi \left( M_i \frac{t_i - \theta}{\sigma} \right) \right). 
\]
Since $\Phi(x)$ is a log concave function, the log-likelihood function has a unique maximizer $\widehat{\theta}_n$ which is the ML estimator.  \\

Next, we show that under the assumptions above, the sequence of messages $M^n$ defines a local asymptotic normal (LAN) family of probability distributions. Therefore, the MSE in estimating $\theta$ from the messages $M^n$ satisfies a local asymptotic minimax property with respect to the \emph{precision} parameter of this LAN family \cite{van2000asymptotic}. Moreover, we conclude the ML estimator is locally asymptotic minimax.

\begin{thm} \label{thm:LAN}
Consider the sequence of threshold detectors $M^n$ with threshold density converges to a probability measure $\lambda$. The the following two statements hold:
\begin{itemize}
\item[(i)] Any estimator ${\theta}_n$ of $\theta$ which is a function of $M^n$ satisfies
\[
\liminf_{c\rightarrow \infty}\, \liminf_{n\rightarrow \infty} \sup_{\tau\,:\,| \tau - \theta| \leq \frac{c}{\sqrt{n}} }  n \mathbb E \left({\theta}_n - \tau \right)^2 \geq \sigma^2/K(\theta),
\]
where 
\[
K(\theta) \triangleq \int \eta\left(\frac{t-\theta}{\sigma} \right) \lambda(dt),
\]
and 
\[
 \eta(x) = \frac{ \phi^2\left( x \right)}{ \Phi\left( x \right) \left(1-\Phi\left(x\right) \right)}. 
\]
\item[(ii)] The asymptotic distribution of the ML estimator $\widehat{\theta}_n$ is given by
\[
\sqrt{n}(\widehat{\theta}_n - \theta) \overset{D}{\rightarrow} \mathcal N\left(0,\sigma^2/K(\theta) \right).
\]
\end{itemize}
\end{thm}

\subsubsection*{Sketch of Proof}
The proof shows that the probability distribution of the vector of messages $M^n$ define a local asymptotic normal (LAN) family of probability distributions with precision parameter $K(\theta)/\sigma^2$. The statements in the theorem then follows from the local asymptotic minimax theorem of LAN families \cite{van2000asymptotic}. The details are given in the Appendix.\\

%Theorem~\ref{thm:dist_thresholds} shows that the maximum likelihood estimator of $\theta$ given the $n$ single-bit messages $M^n$ is asymptotically normal and attains the local asymptotic minimax error $\sigma^2/K(\theta)$.

It follows from Thm.~\ref{thm:LAN} that the asymptotic risk of the ML estimator is $\sigma^2/K(\theta)$, and that this risk is asymptotically minimal over all local alternative estimators for $\theta$. Moreover, the relative efficiency of the ML in the threshold detection scheme equals $1/K(\theta)$. Since the thresholds density $\lambda(t)$ integrates to $1$, and from the bound on the function $\eta(x)$ obtained in the proof of Lemma~\ref{lem:bound_intervals}, it follows that
\[
K(\theta) \leq \sup_{t\in \mathbb R} \eta \left( \frac{t-\theta}{\sigma} \right) \int  \lambda(dt)  \leq \frac{2}{\pi}.
\]
This upper bound on $K(\theta)$ implies that the relative efficiency of any distributed estimator is at least $\pi/2$, a fact that agrees with the lower bound under adaptive estimation derived in Thm.~\ref{thm:adpative_lower_bound}. Unfortunately, this upper bound on $K(\theta)$ is attained whenever the density $\lambda$ is the mass distribution at $\theta$, which is unknown apriori. Therefore, in the section below we consider the asymptotic threshold density function $\lambda$ that minimize the expected asymptotic risk $\mathbb E\left [\sigma^2/K(\theta) \right]$.

{\bf Worst case $\theta$}: in order to bound the local asymptotic minimax risk it is required to consider the desnity $\lambda$ that minimizes the worst case $\theta$ in the parameter space $\Theta$. We assume that $\Theta = [-b,b]$ for some $b>0$, and consider 
\begin{equation} \label{eq:variational}
\sup_{\lambda} \inf_{\theta \in [-b,b]}  {K(\theta)} = \int \eta\left( \frac{t-\theta}{\sigma} \right) \lambda (dt),
\end{equation}
where the suprimum is over all probability measure $\lambda$ on $\mathbb R$. Since \eqref{eq:variational} is concave in $\lambda$, this optimization problem can be solved using a convex program. 
As illustrated in Fig.~\ref{fig:cvx}, the threshold density that maximizes \eqref{eq:variational} is supported over a discrete set in the interval $[-b,b]$. Fig.~\ref{fig:cvx2} illustrates the asymptotic ML risk as a function of the length $2b$ of the parameter space $\theta$.

\begin{figure}
\begin{center}
\begin{tikzpicture}
\node[int] at (0,0)  {\includegraphics[scale=0.2]{support_b20}};
\node[int] at (4.1,0)  {\includegraphics[scale=0.2]{support_b10}};
\node[int] at (0,3.2)  {\includegraphics[scale=0.2]{support_b5}};
\node[int] at (4.1,3.2)  {\includegraphics[scale=0.2]{support_b1}};
\end{tikzpicture}
\caption{\label{fig:cvx}
Support of the optimal density $\lambda$ that minimizes the asymptotic ML risk for the worst choice of $\theta \in [-\sigma b, \sigma b]$ (blue), for $b=1,5,10,20$ (clockwise). The red curve is the value of $K(\theta)$ under this density. 
}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\begin{tikzpicture}
\node at (0,0)  {\includegraphics[scale=0.25]{K_vs_b}};
\end{tikzpicture}
\caption{\label{fig:cvx2}
Asymptotic risk of the ML estimator under an optimal choice of the threshold density $\lambda$ versus the support of the parameter space $\Theta$.
}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\begin{tikzpicture}
\node[int] at (0,0)  {\includegraphics[scale=0.2]{opt_density_sig025}};
\node[int] at (4,0)  {\includegraphics[scale=0.2]{opt_density_sig1}};
\node[int] at (8,0)  {\includegraphics[scale=0.2]{opt_density_sig2}};
\end{tikzpicture}
\caption{\label{fig:opt_density}
The optimal asymptotic threshold density $\lambda$ that minimizes the expected asymptotic ML risk \eqref{eq:cvx_average} for a Gaussian prior for
 $\sigma/\sigma_\theta=0.25,1,2$ (left to right), where $\sigma_\theta^2$ is the variance of the prior. 
}
\end{center}
\end{figure}

\subsection{Average Asymptotic Risk}
Consider problem of minimizing the expected asymptotic risk $\mathbb E[ \sigma^2/K(\theta)]$ over all probability measures $\lambda(dt)$. This optimization problem can be written as follows:
\begin{align}
\label{eq:cvx_average}
\mathrm{minimize} \quad &  \sigma^2 \int \frac{\pi(d\theta)}{K(\theta)} =  \sigma^2 \int \frac{\pi(d\theta)}{ \int \eta \left( \frac{t-\theta}{\sigma}\right) \lambda(dt)}. \\ \nonumber
\mathrm{subject~to} \quad & \lambda(dt)\geq 0,\quad \int \lambda(dt) =1. 
\end{align}
Since the function $x \rightarrow 1/x$ is convex for positive values, \eqref{eq:cvx_average} defines a convex optimization problem in $\lambda$ whose solution depends on the the prior $\pi(\theta)$. The optimal asymptotic threshold density obtained as the solution to \eqref{eq:cvx_average} for a normal prior is illustrated in Fig.~\ref{fig:opt_density}, whereas Fig.~\ref{fig:dist_bound_Gaussian} illustrates the corresponding expected risk. The expected asymptotic risk for the uniform distribution is illustrated in Fig.~\ref{fig:dist_bound_uniform}. 
\par
%
\begin{figure}
\begin{center}
\begin{tikzpicture}
\node at (0,0) {\includegraphics[scale=0.5]{dist_bound_Gauss_prior}};
\node at (0,-2.5) {$\sigma/\sigma_\theta$};
\node[rotate = 90] at (-3.5,0) {(expected asymptotic risk)$/\sigma^2$};
\node at (1,0) {$\pi(\theta) = \mathcal N(0,\sigma_\theta^2)$ };
%\node at (-2,-1.05) {$\frac{\pi}{2}$};
%\node[rotate = -45] at (-1.8,1.2) {\color{red} true };
%\node[rotate = -70] at (-0.5,0) {\color{blue} bound };
\end{tikzpicture}
\caption{Expected asymptotic risk $\mathbb E (1/K(\theta))$ in estimation from threshold detectors versus $\sigma/\sigma_\theta$ for a Gaussian prior with variance $\sigma_\theta^2$. The optimal asymptotic threshold density $\lambda$ (red) is obtained by solving \eqref{eq:cvx_average}. The bound (blue) is obtained using \eqref{eq:upper_bound}.
\label{fig:dist_bound_Gaussian}  }
\end{center}
\end{figure}
%
\begin{figure}
\begin{center}
\begin{tikzpicture}
\node at (0,0) {\includegraphics[scale=0.5]{dist_bound_uniform_prior}};
\node at (0,-2.5) {$\sigma/\sigma_\theta$};
\node[rotate = 90] at (-3.5,0) {(expected asymptotic risk)$/\sigma^2$};
\node at (1,0) {$\pi(\theta)$ = Unif$(- \sqrt{3}\sigma_\theta, \sqrt{3}\sigma_\theta)$ };
%\node at (-2,-1.05) {$\frac{\pi}{2}$};
%\node[rotate = -45] at (-1.8,1.2) {\color{red} true };
%\node[rotate = -70] at (-0.5,0) {\color{blue} bound };
\end{tikzpicture}
\caption{Expected asymptotic risk $\mathbb E (1/K(\theta))$ in estimation from threshold detectors versus $\sigma/\sigma_\theta$ for a uniform prior with variance $\sigma_\theta^2$. The optimal asymptotic threshold density $\lambda$ (red) is obtained by solving \eqref{eq:cvx_average}. The bound (blue) is obtained using \eqref{eq:upper_bound}. 
\label{fig:dist_bound_uniform}  }
\end{center}
\end{figure}

>>>>

%\documentclass[letterpaper]{IEEEtran}
\documentclass[letterpaper, conference]{IEEEtran}      % Use this line for a4 paper
%\IEEEoverridecommandlockouts                              %

@@

\title{\LARGE \bf Mean Estimation from Adaptive Single-bit Measurements}
%
%\author{
%\IEEEauthorblockN{Alon Kipnis}
%\IEEEauthorblockA{Department of Electrical Engineering \\
%Stanford University\\
%Stanford, CA\\}
%\and
%\IEEEauthorblockN{John C. Duchi}
%\IEEEauthorblockA{Department of Statistics \\
%and Department of Electrical Engineering \\
%Stanford University\\
%Stanford, CA\\}
%\and
%\IEEEauthorblockN{Andrea J. Goldsmith}
%\IEEEauthorblockA{Department of Electrical Engineering \\
%Stanford University\\
%Stanford, CA\\}
%}

@@

@@

\node[above of = source, node distance = 1cm] (dist) {$X_i \sim {\mathcal N} \left(\theta, \sigma^2 \right)$};
\node[below of = source2, node distance = 0.5cm] {$\vdots$};

\node[int1, right of = enc2, node distance = 4cm ] (est) {Est};
\draw[->,line width = 2pt] (enc1) -| node[above, xshift = -2cm] (mes1) {$M_1 \in \left\{-1,1\right\}$} (est);   
\draw[->,line width = 2pt] (enc2) -- node[above, xshift = 0cm] (mes2) {$M_2 \in \left\{-1,1\right\}$} (est);

@@

\node[right of = est, node distance = 1.5cm] (dest) {$\widehat{\theta}_n$};
%         \node [int1] (dec) [right of=dest, node distance = 1.5cm,  align=center] {\small Dec };
%\node [int1] (enc) [right of = dec, node distance = 3cm]{Enc}; 
%\draw[->,line width=2pt] (dec) -- (dest);
\draw[->, line width=2pt] (est) -- (dest);
\end{tikzpicture}
\end{center}
\caption{\label{fig:sequential} Distributed single-bit encoding.}
\end{figure}

Consider the distributed 1-bit estimation setting of Fig.~\ref{fig:sequential} where $X_n = Z_n + \theta$ with $Z_n$ zero mean and variance $\sigma^2$. Upon observing the $n$th sample $X_n$, the estimator is observing $M_n(X_n) \in \{0,1\}$, where the mapping $M_n : \mathbb R \rightarrow \{0,1\}$ are determined offline in a manner that will be explained below. Denote $A_n = M_n^{-1}(1)$ and set 
\[
U_n = \begin{cases} 1, & \theta \in A_n \\
-1, & \theta \notin A_n. 
\end{cases}
\]
So the $U_n$s are binary random variables that depends only on the initial draw of $\theta$ and the \emph{detection intervals} $A_1,\ldots,A_n$. We define an \emph{error} event $\mathcal E_n$ at the $n$th observation if $U_n \neq M_n(X_n)$, i.e., an error occurs if $\theta \in A_n$ but $\theta + Z_n \notin A_n$ or $\theta \notin A_n$ but $\theta +Z_n \in A_n$. With this notation, the $n$ observation can be seen as the result of noisy transmission of  $U_n$ through a binary channel with bit flip probability of $\mathbb P( \mathcal E_n)$. We now choose the detection regions $A_n$ as follows: we divide the real line into $m = \lfloor   \sqrt{n} /  (\sigma \alpha_n) \rfloor$ equiprobable disjoint intervals $I_1,\ldots,I_m$. For $j=1,\ldots,m$, include $I_j$ in $A_n$ with probability $1/2$. As a result, we can think of the index of the interval where $\theta$ falls as the \emph{message} (one of possible $m$) that is communicated by $n$ uses of the channel from $U_n$ to $M_n(X_n)$. The width of each interval is $\sigma \alpha_n/ \sqrt{n}$, so if this messages is communicated with zero error, we attain an estimate $\widehat{\theta}$ of accuracy 
\[
\left| \theta - \widehat{\theta} \right|  <  \sigma \frac{\alpha_n}{\sqrt{n}}. 
\]

We now derive conditions of $\alpha_n$ such that communication with vanishing probability of error is possible. Note that by the codeword construction we have
\[
\mathbb P(M_n(X_n)=1) = 1/2,
\]
and, assuming $n$ is large enough,
\[
\mathbb P(\mathcal E_n) = \sum_{j=1}^m \mathbb P(\theta \in I_j)  \mathbb P(X_n \notin I_j) /2 \approx 0.5 \mathbb P( |Z_n|> \sigma \alpha_n/\sqrt{n}) \approx 0.5 -  \frac{1}{\sqrt{2\pi}} \frac{\alpha_n}{\sqrt{n}}+O(1/n).
\]
If $\alpha_n / n \rightarrow 0$, then the probability of bitflip goes to $0.5$ as $n$ increases. By the channel coding theorem we know that communication with vanishing probability is possible provided the communication rate
\[
R = \log_2(m) = \log_2(\sqrt{n} / (\sigma \alpha_n)) 
\]
does not exceeds the capacity:
\[
 C = n(1-h( \mathbb P(\mathcal E))) \approx \frac{n}{2\pi \ln 2} \frac{\alpha_n^2}{n} .
\]
The last condition is satisfied provided 
\[
\alpha_n > \sqrt{2 \pi \ln 2 \log_2(\sqrt{n}/\sigma)}.
\]

{\color{red}
Use Fano's inequality to prove a necessary condition on $\alphan_n$ and therefore on asymptotic relative efficiency.
}

%We use
%\[
%\Phi(x) = 0.5  + \frac{1}{\sqrt{2 \pi} }e^{-x^2/2} \left[x+\frac{x^3}{3}+... \right] 
%\]
%\[
%h(0.5 - x) = 1- \frac{1}{2 \ln 2} \sum_{n=1}^\infty  \frac{2^{2n} x^{2n}} {n(2n-1)}
%\]
